The error in Bayesian phylogenetic reconstruction when speciation co-occurs

### 2019-09-03 razzo-discussion.tex
Improve discussion

For the generative tree prior, we picked the BD tree prior.

 * Choice of generative (JC69 and strict) DNA alignment

 * Choice of generative tree prior (BD)
 * Using Nested Sampling to do model comparison
 * Using the absolute nLTT statistic as an error function

@@

 * MBD parameters
From the four MBD parameters $\lambda$, $\mu$, $\nu$ and $q$,
we investigated 1, 2, 4 and 3 different values respectively.
We chose to use only one $\lambda$, as the proportion of species
created in a co-occurent speciation event is dependent on the ratio
between $\lambda$ and a combination of $\nu$ and $q$.

 * Crown age, and thus the number of taxa in the true species tree
 * Number of replicates
 * DNA sequence length
 * Effective sample size
 * Choice of candidate models

We had to constrain ourselves on the number of replicates.

We had to constrain ourselves on the number of taxa in
the true species trees.

@@

>>>>

When simulating our phylogenies, we used an MBD model with and without
extinction. In the case in which extinction is present,
picking the BD tree prior as the generative tree prior seems justified:
it is the best-fitting standard tree prior. For the MBD model without
extinction, \richel{@Giappo: I think this is an interesting insight:} 
we could have picked the Yule tree prior. We \richel{at least I do} predicted
that the Yule tree prior will be selected as the best candidate model
more often in this scenario, which shows the sensitity of the method we used.

To compare multiple inference models, we picked the Nested Sampling
approach to directly compare marginal likelihoods. We could have used
alternative methods like \richel{forgot the term for now: where you
do an MCMC that can switch between models}. The reason we used Nested
Sampling was mostly pragmatic and we think other methods are equivalent.

We used the absolute nLTT statistic to measure the difference between
two phylogenies. Because the crown age of all phylogenies are equal,
normalization on the time has no effect at all. Also because of these
same crown ages, we could have picked many tree difference metrics, like
\richel{must check:} Pagel's delta or Pigot's delta-r. The reason we used
the nLTT statistic was mostly pragmatic. \verb;pirouette; does allow to 
use a different tree difference metric.

@@

From the four MBD parameters $\lambda$, $\mu$, $\nu$ and $q$,
we investigated 1, 2, 4 and 3 different values respectively.
We chose to use only one $\lambda$, as the proportion of species
created in a co-occurent speciation event is dependent on the ratio
between $\lambda$ and a combination of $\nu$ and $q$.

The crown age we selected for our experiments was based on trial and error
of pilot experiments. Because the number of species in an MBD process is
expected to increase exponentially, increasing the crown age has a
profound effect on the number of taxa, with the consequence of having
runs that took days to calculate. We decided to prefer a lower number of
taxa at the gain of having more replicates. As a consequence, 
this made our research findings 
\richel{that is, I predict, a bigger error in settings with
more co-occurring speciation} more likely to happen: we know that the
choice of a Bayesian prior is most important in weaker data.

The DNA sequence length we used was 1000 base pairs. This is a DNA
sequence length that we \richel{at least I} feel is informative enough
for the amount of taxa we simulate. Additionally (although sequencing
techniques advance at a great pace), this is -to our knowledge- 
of the same magnitude as the DNA length sequenced in the field.

Ideally, our ESSes would all be above the recommended value of 
200 \richel{I predict not all will be above 200, simply due to 
stochasticity}. Also here, we preferred to have more replicates
and a minimum median of 200 over less replicates.

We compare our generative to model to \richel{I predict we will} 
a limited set of candidate models. Also this was for practical
reasons.

@@

The effect of using the absolute nLTT tree statistic to quantify the error
is unknown. A pilot experiment \richel{pirouette example 14} shows that
using the logarithmically transformed nLTT statistic gives a messier and
flatter error distribution, but this is not to be taken for granted.

We chose to use few taxa (as we preferred more replicates), but we do not
know the results would the data be more informative (that is, have more taxa).
A cool feature of \verb;pirouette; is that it can quantify when this happen;
when the prior is uninformative due to strong data.
